An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2016-10-13 aistats2017.tex
updates

We ran MCMC with with our test against the test
from~\citet{icml2014c1_bardenet14} and a simplified version
of~\citet{cutting_mh_2014}. Both methods use repeated testing which
must be discounted to bound overall test
accuracy. ~\citet{icml2014c1_bardenet14} conservatively partitions the
test error, while~\citet{cutting_mh_2014} used a complex dynamic
programming heuristic to estimate the overall test error. We did not
implement this heuristic and instead set the individual test error
equal to the overall test error. Thus our results for~\citet{cutting_mh_2014} should be treated as lower bounds. All
methods were initialized with minibatch size 100.  For~\citet{cutting_mh_2014} we
increment sizes by 100 and set the tolerance $\epsilon=0.005$.  For~\citet{icml2014c1_bardenet14}, we increase sizes geometrically with $\gamma = 1.5$ and
use parameters $p = 2, \delta = 0.01$.  All methods collect 5000
samples using a random walk proposer with covariance ${\rm diag}(0.15,
0.15)$, which means the M-H test is responsible for shaping the sample
distribution.

@@

We next test logistic regression for the binary classification of 1s
versus 7s on a subset of the MNIST8M dataset, a larger version of
MNIST~\citep{lecun-mnisthandwrittendigit-2010}. We randomly subsampled
450k training and 192k testing points.
%%% The pixels are scaled in $[0,1]$  - JFC Why ????
We impose a uniform prior on $\theta$ and again use a random walk
proposer, this time with covariance matrix $0.05I$. The temperature is
set at $T=1000$. We run the three tests for 3000 samples and again use 100
as the starting minibatch size. Method~\citet{cutting_mh_2014} used
per-test error of $\epsilon=0.05$.  For~\citet{icml2014c1_bardenet14},
we use their recommended discounting factor value of $\gamma=2.0$. We tried to use the
symbolic bound for $C_{\theta,\theta'}$ for Logistic regression
from~\citet{icml2014c1_bardenet14}, but found it to be too high. Instead we use
the empirical $C_{\theta,\theta'}$ from the entire dataset and do not add its
computation time when considering runtime, so again our results represent a
lower bound for this method.

@@

Figure~\ref{fig:logistic_minibatch} shows histograms of minibatch sizes for the
three methods on a log-log scale. With an initial size of 100, we achieve an
average minibatch size of 393, far smaller than the averages of the other two
methods, which have longer-tailed distributions.

>>>>

We ran MCMC with with our test against the test
from~\citet{icml2014c1_bardenet14} and a simplified version
of~\citet{cutting_mh_2014}. Both methods use repeated testing which
must be discounted to bound overall test
accuracy. ~\citet{icml2014c1_bardenet14} conservatively partitions the
test error, while~\citet{cutting_mh_2014} used a complex dynamic
programming heuristic to estimate the overall test error. We did not
implement this heuristic and instead set the individual test error
equal to the overall test error. Thus our results
for~\citet{cutting_mh_2014} should be treated as lower bounds. All
methods were initialized with minibatch size 100.
For~\citet{cutting_mh_2014} we increment sizes by 100 and set the
tolerance $\epsilon=0.005$ to control overall test error.
For~\citet{icml2014c1_bardenet14}, we increase sizes geometrically
with $\gamma = 1.5$ and use parameters $p = 2, \delta = 0.01$.  All
methods collect 5000 samples using a random walk proposer with
covariance ${\rm diag}(0.15, 0.15)$, which means the M-H test is
responsible for shaping the sample distribution.

@@

We next test logistic regression for the binary classification of 1s
versus 7s on a subset of the MNIST8M dataset, a larger version of
MNIST~\citep{lecun-mnisthandwrittendigit-2010}. We randomly subsampled
450k training and 192k testing points.
%%% The pixels are scaled in $[0,1]$  - JFC Why ????
We impose a uniform prior on $\theta$ and again use a random walk
proposer, this time with covariance matrix $0.05I$. The temperature is
set at $T=1000$. We run the three tests for 3000 samples and again use 100
as the starting minibatch size. Our undiscounted implementation of~\citet{cutting_mh_2014} used
per-test error of $\epsilon=0.05$.  For~\citet{icml2014c1_bardenet14},
we use their recommended discounting factor value of $\gamma=2.0$. We tried to use the
symbolic bound for $C_{\theta,\theta'}$ for Logistic regression
from~\citet{icml2014c1_bardenet14}, but found it to be too high. Instead we use
the empirical $C_{\theta,\theta'}$ from the entire dataset and do not add its
computation time when considering runtime, so again our results represent a
lower bound for this method.

@@

Figure~\ref{fig:logistic_minibatch} shows histograms of minibatch sizes for the
three methods on a log-log scale. With an initial size of 100, we achieve an
average minibatch size of 393, about two orders of magnitude smaller than
~\citet{icml2014c1_bardenet14}, and an order of magnitude faster than
the undiscounted implementation of~\citet{cutting_mh_2014}.
